---
title: "Day Nine"
author: "Kelly Bodwin"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(here)
library(janitor)
```

## Part One

```{r}
#dat <- readLines(here("09-example"))
dat <- readLines(here("09-data"))

mat <- dat %>%
  map(~str_split(.x, "") %>% unlist() %>% as.numeric()) %>%
  unlist() %>%
  matrix(ncol = length(dat)) %>%
  t()

n_rows = length(dat)
n_cols = ncol(mat)
```

```{r}
lower_than_right <- mat - cbind(mat[, -1], rep(10, n_rows)) < 0

lower_than_left <- mat - cbind(rep(10, n_rows), mat[, -n_cols]) < 0

lower_than_above <- mat - rbind(mat[-1,  ], rep(10, n_cols)) < 0

lower_than_below <- mat - rbind(rep(10, n_cols), mat[-n_rows,]) < 0
```

```{r}
all <- lower_than_left & lower_than_right & lower_than_above & lower_than_below

sum(mat[all] + 1)
```
## Part Two

Okay, so a basin always flows to a low point, which we've already found.

I've already found the flows "lower than adjacent"

What I need to know is how many TRUEs in a row there are up to the low point.

... but not quite in a row, because they can flow down and left etc.  I need to know the size of the adjacent "blob" of TRUEs.

Fucking recursion i hate it.

```{r}

flow_down <- function(point) {
  
  if (mat[point[1], point[2]] > 8) {
    return(NULL)
  }
  
  adjacents <- c(
    right = mat[point[1], point[2] + 1],
    left = mat[point[1], point[2] - 1],
    up = mat[point[1] - 1, point[2]],
    down = mat[point[1] + 1, point[2]]
  )
  
  adjacents <- adjacents[adjacents < mat[point[1], point[2]]]
  
  while (length(adjacents) > 0) {
    
    thing <- names(adjacents)[which.min(adjacents)]
    
    if (thing == "up") {
      
      point <- flow_down(c(point[1] - 1, point[2]))
      
    } else if (thing == "down") {
      
      point <- flow_down(c(point[1] + 1, point[2]))
      
    } else if (thing == "left") {
      
      point <- flow_down(c(point[1], point[2] - 1))
      
    } else if (thing == "right") {
      
      point <- flow_down(c(point[1], point[2] + 1))
      
    }
    
     adjacents <- c(
    right = mat[point[1], point[2] + 1],
    left = mat[point[1], point[2] - 1],
    up = mat[point[1] - 1, point[2]],
    down = mat[point[1] + 1, point[2]]
  )
  
  adjacents <- adjacents[adjacents < mat[point[1], point[2]]]
    
  }
  
  return(point)
  
}

flow_down(c(1,1))

```

```{r}
## matrix needs a stupid boundary

mat <- cbind(mat, rep(10, 100))
mat <- rbind(mat, rep(10, 101))

all_locations <- expand.grid(1:100, 1:100)

basin_lows <- map(1:nrow(all_locations),  ~flow_down(c(all_locations[.x,1], 
                                                       all_locations[.x, 2])))

basin_sizes <- 
  basin_lows %>%
  map_chr(~str_c(.x, collapse = "")) %>%
  table() %>%
  sort()
```

arg 9's don't count as basins, that's so rude.  Fixed it in the function.

```{r}
basin_sizes %>%
  rev()

96*92*89
```

